### 2025-05-28 chapters/mpif/ms.Rnw
Fixed references

An extensively used approach in ecology, known as \emph{data cloning} \citep{lele07,lele10}, involves iteratively using Bayes' rule, by recursively mapping a prior distribution to a posterior distribution using the same likelihood function, until the iterated posterior distribution converges to a point mass at the maximum likelihood estimate (MLE).
Data cloning is equivalent to performing a Bayesian update on multiple copies of the data when the data are independent and identically distributed.
\citet{ionides15} developed an extension of data cloning for nonlinear non-Gaussian state space models by treating the model parameters as latent states, and performing a random walk for these parameters at each observation time.
By iteratively applying a particle filter \citep{arulampalam02} to this extended model and decreasing the random walk standard deviations over time, it can be shown that the parameters will converge to the MLE after a sufficiently large number of iterations \citep{ionides15, chen24}.
This algorithm, known as IF2, has seen extensive use for inference on low dimensional dynamic models, especially in epidemiological contexts \citep[e.g.,][]{pons-salort18,stocks18,subramanian21,fox22}.

>>>>

An extensively used approach in ecology, known as \emph{data cloning} \citep{lele07,lele10}, involves iteratively using Bayes' rule, by recursively mapping a prior distribution to a posterior distribution using the same likelihood function, until the iterated posterior distribution converges to a point mass at the maximum likelihood estimate (MLE).
Data cloning is equivalent to performing a Bayesian update on multiple copies of the data when the data are independent and identically distributed.
\citet{ionides15} developed an extension of data cloning for nonlinear non-Gaussian state space models by treating the model parameters as latent states, and performing a random walk for these parameters at each observation time.
By iteratively applying a particle filter \citep{arulampalam02} to this extended model and decreasing the random walk standard deviations over time, it can be shown that the parameters will converge to the MLE after a sufficiently large number of iterations \citep{ionides15, chen24}.
This algorithm, known as IF2, has seen extensive use for inference on low dimensional dynamic models, especially in epidemiological contexts \citep[e.g.,][]{pons-salort18,stocks20,subramanian21,fox22}.
